Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-22 paper/current/main.tex
update text relating to vector length plot

%
% Figure - SemAxis results
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Length of organization vectors captures a connection diversity of organizations.}
		\textbf{a} L2 norm of organization vectors compared to their number of researchers for U.S. universities.
		The L2 norm corresponds to the distance from the origin of the vector space.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest rank university.
		Uncolored points are universities not listed on a times rank.
		A "boomerang"-shape emerges, wherein larger universities tend to more distant from the origin (large L2 norm), however the more prestigious the university, the closer it is to the origin (small L2 norm). 
		\textbf{b} This boomerang pattern repeats globally for all organizations across all countries, shown for the United States, Germany, China, and others (loess regression line shown). 
		The L2 norm is correlated weakly with the number of connections divided by the number of researchers in the organization, which can be interpreted diversity of the connection compared to their size ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit.

	}
	\label{fig:semaxis}
\end{figure}

@@

%
% Figure - s_i by factors
%
\begin{figure}[hp!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/si_factors.pdf}
	\caption{
		\textbf{Factors relating to $s_{i}$ among universities}
		Correlation between various organization-level factors for U.S. universities and the value $s_{i}$.
		Dots correspond to organizations.	
		The red line is the line of  the best fit. 
		Red text is the regression estimate.
		Number of authors is the average annual count of unique mobile and non-mobile authors.
		Rankings are derived from the Times Ranking of World Universities, and the Leiden Rankings of Universities. 
		Remaining variables come from the Carnegie Classification of Higher Education Institutions.
		The factors that best explain $s_i$ are the number of authors, the rank, the amount of Science and Engineering (S\&E) funding, and the number of doctorates granted. 
	}
	\label{fig:supp:semaxis_states}
\end{figure}

@@

>>>>

What is it about the organization embedding that captures prestige so effectively?
We find that the length of organization vectors---that is, the distance from the origin of the vector space---provides provides insight into the characteristics of organizations. 
Up to a point, the length of U.S. organization's vectors tends to increase proportional to its number of researchers (both mobile and non-mobile) who published with that organization;
these organizations are primarily teaching-focused institutions, agencies, and hospitals that either are not ranked or have a low ranking. 
However, at around 1,000 researchers, the length of the vector decreases with increase in the number of researchers;
organizations here are primarily research-intensive and prestigious universities with higher rankings, research outputs, R\&D funding, and doctoral students (Fig.~\ref{fig:supp:length_vs_metaInfo}).
This same convex-curve repeats irrespective of country, such as China and Germany, though the curve may be shifted based on the typical size and vector length of their organizations (Fig.~\ref{fig:length}b).

This pattern has been observed in applications of neural embedding to natural language, wherein the length of word vectors were found to represent word "significance"~\autocite{schakel2015measuring
Here, interpret the vector length instead as a measure of an organization's connectedness to other organizations given its size, where a connection is defined as a non-zero flow between organizations.
for example, organizations with a small L2 norm tend to also have a lower number of connections given their size (Fig.~\ref{fig:length}b, inset, $R^{2} = 0.17$). 
Large and prestigious universities can be considered as having a shorter vector length than expected given their size, reflecting their ability to attract researchers from many places, to place graduates in positions around the world, and the capacity to encourage mobility by offering temporary and visiting researcher positions. 
At the national level, the relative position of the convex-curves can reveal characteristics of their scientific organizations—for example, compared to the United States, Chinese organizations have a longer typical vector length (Fig.~\ref{fig:length}b) reflecting lower overall connectedness in the embedding space. 
Prestige is a key component academic mobility—by encoding information relating to prestige, neural embeddings offer a means of better representing functional distance between organizations and capturing the global landscape of mobility.

%
% Figure - SemAxis results
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Length of organization vectors captures connection diversity of organizations.}
		\textbf{a.} L2 norm of organization vectors compared to their number of researchers for U.S. universities.
		The L2 norm corresponds to the distance from the origin of the vector space.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest ranked university.
		Uncolored points are universities not listed on a Times ranking.
		A convex-shape emerges, wherein larger universities tend to more distant from the origin (large L2 norm), however the more prestigious the university, the closer it is to the origin (small L2 norm). 
		\textbf{b.} This convex pattern repeats globally for all organizations across all countries, shown for the United States, Germany, China, and others (loess regression line shown). 
		The L2 norm is correlated weakly with the number of connection divided by the number of researchers in organization, which can be interpreted diversity of the connection compared to their size ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit.
	}
	\label{fig:length}
\end{figure}

@@

@@

%
% Figure - length by factors
%
\begin{figure}[hp!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/si_factors.pdf}
	\caption{
		\textbf{Factors relating to the L2 norm of vectors for U.S. universities}
		Correlation between the L2 norm of organization embedding vectors of U.S. universities and characteristics of U.S. universities. 
		Dots correspond to organizations.	
		The red line is the line of  the best fit. 
		Red text is the regression estimate.
		Number of authors is the average annual count of unique mobile and non-mobile authors.
		Rankings are derived from the Times Ranking of World Universities, and the Leiden Rankings of Universities. 
		Remaining variables come from the Carnegie Classification of Higher Education Institutions.
		The factors that best explain $s_i$ are the number of authors, the rank, the amount of Science and Engineering (S\&E) funding, and the number of doctorates granted. 
	}
	\label{fig:supp:length_vs_metaInfo}
\end{figure}
